---
title: "Body Fat Prediction"
author: "GS"
date: "18/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, comment = "", warning = F, fig.align = "center")
```

#### Context

Lists estimates of the percentage of body fat determined by underwater weighing and various body circumference measurements for 252 men.

#### Content

- The variables listed below, from left to right, are:
- Density determined from underwater weighing
- Percent body fat from Siri's (1956) equation
- Age (years)
- Weight (lbs)
- Height (inches)
- Neck circumference (cm)
- Chest circumference (cm)
- Abdomen 2 circumference (cm)
- Hip circumference (cm)
- Thigh circumference (cm)
- Knee circumference (cm)
- Ankle circumference (cm)
- Biceps (extended) circumference (cm)
- Forearm circumference (cm)
- Wrist circumference (cm)

(Measurement standards are apparently those listed in Benhke and Wilmore (1974), pp. 45-48 where, for instance, the abdomen 2 circumference is measured "laterally, at the level of the iliac crests, and anteriorly, at the umbilicus".)

These data are used to produce the predictive equations for lean body weight given in the abstract "Generalized body composition prediction equation for men using simple measurement techniques", K.W. Penrose, A.G. Nelson, A.G. Fisher, FACSM, Human Performance Research Center, Brigham Young University, Provo, Utah 84602 as listed in Medicine and Science in Sports and Exercise, vol. 17, no. 2, April 1985, p. 189. (The predictive equation were obtained from the first 143 of the 252 cases that are listed below).


[Link Kaggle](https://www.kaggle.com/fedesoriano/body-fat-prediction-dataset)

```{r, packages}
library(tidyverse)
library(tidymodels)
library(GGally)
library(corrr)
library(ggcorrplot)
library(ggExtra) # marginal plots
library(gridExtra)
library(wesanderson)

theme_set(theme_bw())
```

```{r, dataset}
fat <- read_csv("bodyfat.csv")
```
### EDA

```{r, str}
sample_n(fat,5)
dim(fat)
```

**Response variable: BodyFat**

```{r, summary}
summary(fat)
# No NAs
```


```{r,lbs_to_metric}
fat <- fat %>% 
  mutate(Weight = round(Weight * 0.453592, digits = 2),
         Height = round(Height * 2.54, digits = 2))
```

```{r, cache = T}
ggpairs(fat) 
```

```{r, outliers detection, echo = F}
fat %>% 
  filter(BodyFat == 0)
  
fat %>% 
  filter(Height == 74.93)

density <- ggplot(fat, aes(Density)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Density variable",
       x = "Density",
       y = "Density")

p1 <- ggMarginal(density, type = "boxplot", margins = "x")


age <- ggplot(fat, aes(Age)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Age variable",
       x = "Age",
       y = "Density")

p3 <- ggMarginal(age, type = "boxplot", margins = "x")


weight <- ggplot(fat, aes(Weight)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Weight variable",
       x = "Age",
       y = "Density")

p4 <- ggMarginal(weight, type = "boxplot", margins = "x")

height <- ggplot(fat, aes(Height)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Height variable",
       x = "Height",
       y = "Density")

p5 <- ggMarginal(height, type = "boxplot", margins = "x")


neck <- ggplot(fat, aes(Neck)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Neck variable",
       x = "Neck",
       y = "Density")

p6 <- ggMarginal(neck, type = "boxplot", margins = "x")


chest <- ggplot(fat, aes(Chest)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Chest variable",
       x = "Chest",
       y = "Density")

p7 <- ggMarginal(chest, type = "boxplot", margins = "x")


abdomen <- ggplot(fat, aes(Abdomen)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Abdomen variable",
       x = "Abdomen",
       y = "Density")

p8 <- ggMarginal(abdomen, type = "boxplot", margins = "x")


hip <- ggplot(fat, aes(Hip)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Hip variable",
       x = "Hip",
       y = "Density")

p9 <- ggMarginal(hip, type = "boxplot", margins = "x")


thigh <- ggplot(fat, aes(Thigh)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Thigh variable",
       x = "Hip",
       y = "Density")

p10 <- ggMarginal(thigh, type = "boxplot", margins = "x")


knee <- ggplot(fat, aes(Knee)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Knee variable",
       x = "Hip",
       y = "Density")

p11 <- ggMarginal(knee, type = "boxplot", margins = "x")


ankle <- ggplot(fat, aes(Ankle)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Ankle variable",
       x = "Hip",
       y = "Density")

p12 <- ggMarginal(ankle, type = "boxplot", margins = "x")


biceps <- ggplot(fat, aes(Biceps)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Biceps variable",
       x = "Hip",
       y = "Density")

p13 <- ggMarginal(biceps, type = "boxplot", margins = "x")


forearm <- ggplot(fat, aes(Forearm)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Forearm variable",
       x = "Forearm",
       y = "Density")

p14 <- ggMarginal(forearm, type = "boxplot", margins = "x")


wrist <- ggplot(fat, aes(Wrist)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of Wrist variable",
       x = "Wrist",
       y = "Density")

p15 <- ggMarginal(wrist, type = "boxplot", margins = "x")


grid.arrange(p1, p3, p4, p5)
grid.arrange(p6, p7, p8, p9)
grid.arrange(p10, p11, p12, p13)
grid.arrange(p14, p15)

# Response variable
response <- ggplot(fat, aes(BodyFat)) + 
  geom_point(aes(y = 0.01), alpha = 0) + # solo para ggMarginal
  geom_histogram(aes(y = ..density..), bins = sqrt(nrow(fat)), fill = "lightblue", color = "blue") + 
  geom_density() + 
  geom_rug(color = "blue") + 
  labs(title = "Distribution of response variable",
       subtitle = "BodyFat",
       x = "Body Fat",
       y = "Density")

ggMarginal(response, type = "boxplot", margins = "x")
```

```{r, normality}
shapiro.test(fat$BodyFat) # normally distributed
  
p_val <-  NULL
for(i in 1:ncol(fat)){
  p_val[i] = map(fat, shapiro.test)[[i]]$p.value
}

p_val <- tibble(variable = colnames(fat),
                p_val = round(p_val, 4),
                distribution = if_else(p_val >= 0.05, "Normal", "Non normal"))

p_val
```


```{r, outlier removal}
fat <- fat %>%
    filter(if_all(where(is.numeric), ~ (.>= (quantile(., .25) - IQR(.) * 1.5 )) &
           (.<= (quantile(., .75) + IQR(.) * 1.5 ))))


#One implausibel value is BodyFat = 0 which it was not detetc as outlier by IQR*1.5 method
# I delete manually

fat <- fat %>% 
  filter(BodyFat >= 2)

summary(fat)
```


```{r, normality_wo_outliers}
shapiro.test(fat$BodyFat) # normally distributed
  
pp_val <-  NULL
for(i in 1:ncol(fat)){
  pp_val[i] = map(fat, shapiro.test)[[i]]$p.value
}

pp_val <- tibble(variable = colnames(fat),
                pp_val = round(pp_val, 4),
                distribution = if_else(pp_val >= 0.05, "Normal", "Non normal"))

pp_val
```
```{r, uni_analysis}
r.sq <- fat %>% 
  dplyr::select(-BodyFat) %>%  # exclude outcome, leave only predictors 
  map(~lm(fat$BodyFat ~ .x, data = fat)) %>% 
  map(summary) %>%
  map_dbl("r.squared")


p.val <- fat %>% 
  dplyr::select(-BodyFat) %>%  # exclude outcome, leave only predictors 
  map(~lm(fat$BodyFat ~ .x, data = fat)) %>% 
  map(summary) %>% 
  map(c("coefficients")) %>% 
  map_dbl(8)


estimate <- fat %>% 
  dplyr::select(-BodyFat) %>%  # exclude outcome, leave only predictors 
  map(~lm(fat$BodyFat ~ .x, data = fat)) %>% 
  map(summary) %>% 
  map(c("coefficients")) %>% 
  map_dbl(2)  


l_ci <- fat %>% 
  dplyr::select(-BodyFat) %>%  # exclude outcome, leave only predictors 
  map(~lm(fat$BodyFat ~ .x, data = fat)) %>% 
  map(confint) %>% 
  map_dbl(2)


u_ci <- fat %>% 
  dplyr::select(-BodyFat) %>%  # exclude outcome, leave only predictors 
  map(~lm(fat$BodyFat ~ .x, data = fat)) %>% 
  map(confint) %>% 
  map_dbl(4)

vars <- names(fat[,c(1,3,4,5,6,7,8,9,10,11,12,13,14,15)])

plot_fat <- data.frame(vars, estimate, l_ci, u_ci, p.val)
row.names(plot_fat) <- NULL
plot_fat$vars <- as.factor(plot_fat$vars)

knitr::kable(plot_fat)
plot_fat

confint_plt <- plot_fat %>%
  filter(vars != "Density") %>% 
  ggplot(aes(x = vars, y = estimate)) +
  geom_errorbar(aes(ymax = u_ci, ymin = l_ci), color = wes_palette("Rushmore1")[3]) +
  geom_point(color = wes_palette("Royal1")[2], size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  theme_minimal() +
  coord_flip() +
  labs(title = "Coefficients from Univariate Linear Models",
       subtitle = "Density was exclude due to scale",
       x = "",
       y = "")  
confint_plt
```

```{r, correlation}
(corel <- cor(fat) %>% round(., 2))

p.mat <- cor_pmat(fat)
ggcorrplot(corel,
           method = "square",
           type = "lower",
           lab = T,
           lab_size = 3,
           p.mat = p.mat,
           hc.order = T,
           colors = c("#6D9EC1", "white", "#E46726"))

cor.test(fat$Weight, fat$Abdomen)

correlate(fat) %>% 
  network_plot()
```

```{r, cluster analysis}
fat_norm <- scale(fat)

distance <- dist(fat_norm, method = "euclidean")
cluster <- hclust(distance, method = "complete")

plot(cluster)
```
```{r, pca}

library(factoextra)
pca <- prcomp(fat, scale. = T)
pca
summary(pca)

fviz_eig(pca, addlabels = T)

fviz_pca_var(pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

```

## Initals models (exploratory)

```{r, raw_lineal_model}
raw_lineal_model <- lm(BodyFat ~ ., data = fat)
summary(raw_lineal_model)
plot(raw_lineal_model)
```
```{r, multicol}
car::vif(raw_lineal_model)
# I will remove values above 5
```

```{r, feauture_selection_subset}
library(leaps)

subset_lineal_model <- regsubsets(BodyFat ~ ., data = fat, nvmax = 14)
summary(subset_lineal_model)
```


```{r, feauture_selection_subset_metrics}
res.sum <- summary(subset_lineal_model)
tibble(Adj.R2 = which.max(res.sum$adjr2),
           CP = which.min(res.sum$cp),
           BIC = which.min(res.sum$bic)
           )
```

```{r, feature_selection_stepwise}
library(MASS)

stepwise_lineal_model <- stepAIC(raw_lineal_model, direction = "both", trace = F)
summary(stepwise_lineal_model)
```

```{r, splitting}
set.seed(2021)
split_fat <- initial_split(fat, prop = .75, strata = BodyFat)
fat_train <- training(split_fat)
fat_test <- testing(split_fat)

```
```{r, cv}
set.seed(2021)
fat_cv <- vfold_cv(fat_train, v = 10)

```

Basic Linear Rregression

```{r, lm_basic}
lm_model <- linear_reg() %>% 
  set_engine("lm")

lm_basic_rec <- 
  recipe(BodyFat ~ ., data = fat_train )

lm_basic_wf <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(lm_basic_rec)
```

```{r, fit_lm_basic}
set.seed(2021)
fit_lm_basic <- fit_resamples(
  lm_basic_wf,
  resamples = fat_cv,
  control = control_resamples(save_pred = T)
  )
```
```{r, pred_lm_basic}
fit_lm_basic$.predictions
fit_lm_basic %>% collect_metrics()
fit_lm_basic %>% augment() 
```

Preprocessed Linear Model

```{r}
lm_prepro_rec <- 
  recipe(BodyFat ~ ., data = fat_train) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_lincomb(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors(), num_comp = 4)

lm_prepro_rec %>% prep() %>% juice()

lm_prepro_wf <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(lm_prepro_rec)
```
```{r}
set.seed(2021)
fit_lm_prepro <- fit_resamples(
  lm_prepro_wf,
  resamples = fat_cv,
  control = control_resamples(save_pred = T)
)

```
```{r, pred_lm_prepro}
fit_lm_prepro$.predictions
fit_lm_prepro %>% collect_metrics()
fit_lm_prepro %>% augment() 
```




Penalized Model

```{r, penalized model}
penal_model <-
  linear_reg(penalty = tune(), mixture = tune()) %>%
  set_engine('glmnet')

penal_rec <- lm_basic_rec

penal_wf <- 
  workflow() %>% 
  add_model(penal_model) %>% 
  add_recipe(penal_rec)

penal_wf %>% parameters()
penal_wf %>% parameters() %>% pull_dials_object("penalty")
penal_wf %>% parameters() %>% pull_dials_object("mixture")
```
```{r, cache= TRUE}
reg_metric <- metric_set(rsq, rmse)
keep_pred <- control_resamples(save_pred = TRUE, save_workflow = TRUE)

set.seed(2021)

penal_tuning <- 
  tune_grid(
    penal_wf,
    resamples = fat_cv,
    grid = 20,
    metrics = reg_metric,
    control = keep_pred
  )

penal_tuning
```

```{r, penal_metrics}
penal_tuning %>% collect_metrics()
penal_tuning %>% collect_predictions()
penal_tuning %>% show_best(metric = "rmse")
penal_tuning %>% show_best(metric = "rsq")
penal_tuning %>% select_best(metric = "rmse")

autoplot(penal_tuning) + theme(legend.position = "top")
```

```{r, penal_final}
best_penal_model <- 
  penal_tuning %>% 
  select_best(metric = "rmse")

last_penal_wf <- 
  penal_wf %>% 
  finalize_workflow(best_penal_model)

last_penal_wf
```
```{r applying last model}
penal_fit <- 
  last_fit(
    last_penal_wf,
    split = split_fat
  )

penal_fit
```
```{r penal fitted metrics}
penal_fit %>% collect_metrics()
penal_pred <- penal_fit %>% collect_predictions()
penal_pred
```













